---
title: "README"
date: 
output: md_document
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1000)
```


# treeSimR

An R package for easy forward simulating probability decision trees, calculating cost-effectiveness and PSA.

Currently contains functions to:

* read-in and check tree object
* simulate final outcomes
* Monte-Carlo simulate multiple simulations

TODO

* optimal decision function
* plotting functions: C-E plane, C-E curve (others)


## Read-in trees

```{r load packages}
library(yaml)
library(data.tree)
devtools::load_all(".")
```


The raw decision tree file is a tab-spaced file such as the following:

```{r eval=FALSE}
sink("raw data/LTBI_dtree-cost-distns.yaml")

cat("
    name: LTBI screening cost
    type: decision
    distn: gamma
    mean: 1
    sd: 1
    Screening:
      distn: gamma
      mean: 1
      sd: 1
      type: chance
      LTBI:
        p: 0.25
        distn: gamma
        mean: 1
        sd: 1
        type: chance
        Not GP registered:
          type: terminal
          p: 0.4
          distn: gamma
          mean: 1
          sd: 1
        GP registered:
          type: chance
          p: 0.4
          distn: gamma
          mean: 1
          sd: 1
          Not Agree to Screen:
            p: 0.6
            type: terminal
            distn: gamma
            mean: 1
            sd: 1
          Agree to Screen:
            p: 0.6
            type: chance
            distn: gamma
            mean: 1
            sd: 1
            Test Negative:
              type: terminal
              p: 0.7
              distn: gamma
              mean: 1
              sd: 1
            Test Positive:
              type: chance
              p: 0.7
              distn: gamma
              mean: 1
              sd: 1
              Not Start Treatment:
                type: terminal
                p: 0.3
                distn: gamma
                mean: 1
                sd: 1
              Start Treatment:
                type: chance
                p: 0.3
                distn: gamma
                mean: 1
                sd: 1
                Complete Treatment:
                  type: terminal
                  p: 0.75
                  distn: gamma
                  mean: 1
                  sd: 1
                Not Complete Treatment:
                  type: terminal
                  p: 0.75
                  distn: gamma
                  mean: 1
                  sd: 1
      non-LTBI:
        p: 0.25
        distn: gamma
        mean: 1
        sd: 1
        type: chance
        Not GP registered:
          type: terminal
          p: 0.4
          distn: gamma
          mean: 1
          sd: 1
        GP registered:
          type: terminal
          p: 0.4
          distn: gamma
          mean: 1
          sd: 1
          Not Agree to Screen:
            p: 0.6
            type: chance
            distn: gamma
            mean: 1
            sd: 1
          Agree to Screen:
            p: 0.6
            type: chance
            distn: gamma
            mean: 1
            sd: 1
            Test Negative:
              type: terminal
              p: 0.7
              distn: gamma
              mean: 1
              sd: 1
            Test Positive:
              type: terminal
              p: 0.7
              distn: gamma
              mean: 1
              sd: 1
              Not Start Treatment:
                type: terminal
                p: 0.3
                distn: gamma
                mean: 1
                sd: 1
              Start Treatment:
                type: terminal
                p: 0.3
                distn: gamma
                mean: 1
                sd: 1
                Complete Treatment:
                  type: terminal
                  p: 0.75
                  distn: gamma
                  mean: 1
                  sd: 1
                Not Complete Treatment:
                  type: terminal
                  p: 0.75
                  distn: gamma
                  mean: 1
                  sd: 1
    No Screening:
      p: 0.25
      distn: gamma
      mean: 1
      sd: 1
      type: chance
      LTBI:
        type: terminal
        p: 0.4
        distn: gamma
        mean: 1
        sd: 1
      non-LTBI:
        p: 0.6
        type: terminal
        distn: gamma
        mean: 1
        sd: 1
    ", fill=TRUE)
sink()
```


We save this to a .yaml text file and then read it in as a yaml file to a data.tree object using the yaml and data.tree packages. This is then represented as a list in R.

```{r}
# osList <- yaml.load(yaml)
osList <- yaml.load_file("raw data/LTBI_dtree-cost-distns.yaml")
osNode <- as.Node(osList)
osNode
```

Better still, use the package function to do this, checking for tree integrity and defining an additional costeffectiveness.tree class.

```{r}
osNode <- treeSimR::create.costeffectiveness.tree(yaml_tree = "raw data/LTBI_dtree-cost-distns.yaml")
print(osNode, "type", "p", "distn", "mean", "sd")
```



## Simulate a scenario

We can now sample values for each branch, given the distributions defined for each.
This could be the cost or health detrminent.

```{r}
rpayoff <- osNode$Get(sampleNode)
osNode$Set(payoff = rpayoff)
print(osNode, "type", "p", "distn", "mean", "sd", "payoff")
```

Now given the sampled values, e.g. cost, and the probabilities, we can calculate the expected values at each not from leaf to root.

```{r}
osNode$Do(payoff, traversal = "post-order", filterFun = isNotLeaf)

print(osNode, "type", "p", "distn", "mean", "sd", "payoff")
```

Similarly to above, we have created a better wrapper function to perform these steps:

```{r}
osNode <- calc.expectedValues(osNode)
print(osNode, "type", "p", "distn", "mean", "sd", "payoff")
```



## Monte Carlo forward simulation


We are now in a position to do a probability sensitivity analysis (PSA) and calculate multiple realisation for specific nodes e.g. those at which a decision is to be made.

```{r}
MonteCarlo.expectedValues(osNode, n=100)
```


## Optimal decisions

We can get the software to calculate the optimal decision for us, rather than returning the expections to compare

```{r eval=FALSE}
##TODO##
osNode$Do(decision, filterFun = function(x) x$type == 'decision')
osNode$Get('decision')[1]
```


```{r eval=FALSE}
##TODO##
## probabilty of successfully & correctly treating LTBI
dummy <- rep(0, osNode$totalCount)
dummy[12] <- 1
osNode$Set(payoff = dummy)
print(osNode, "type", "p", "distn", "mean", "sd", "payoff")
osNode$Do(payoff, traversal = "post-order", filterFun = isNotLeaf)
print(osNode, "type", "p", "distn", "mean", "sd", "payoff")
osNode$Get('payoff')[1]
```

