---
title: "README"
author: "Nathan Green"
date: "`r Sys.Date()`"
output:
    github_document
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1000)
```


# `treeSimR`

An R package for easy, robust forward simulating probability decision trees, calculating cost-effectiveness and probability sensitivity analysis (PSA).

Currently contains functions to:

* read-in and check tree object
* simulate final expected outcomes
* Monte-Carlo simulate multiple simulations

_TODO_

- [ ] iteratively collapse expected outcome (from right to left)
- [ ] iteratively collapse chance nodes (from right to left)
- [ ] optimal decision function (iterative from right to left)


The package leans heavily on the `data.tree` package,
(introduction [here](https://cran.r-project.org/web/packages/data.tree/vignettes/data.tree.html) and examples [here](https://cran.r-project.org/web/packages/data.tree/vignettes/applications.html) ).



## Installing `treeSimR`

To install the development version from github:
```{r eval=FALSE}
library(devtools)
install_github("n8thangreen/treeSimR")
```

Then, to load the package, use:
```{r eval=FALSE}
library("treeSimR")
```

See package [Vignette](http://htmlpreview.github.io/?https://github.com/n8thangreen/treeSimR/blob/master/inst/doc/vignette_main.html) for more details and examples.

For plotting functions (e.g. C-E plane, C-E curve, risk profile (with uncertainty), tornado, spider, ...) see the [plotCostEffectiveness package](https://github.com/n8thangreen/plotCostEffectiveness).

